Towards Verifying the Bitcoin-S Library

### 2019-07-04 paper/paper.tex
further reduce code

\begin{figure}
\begin{lstlisting}[style=scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}
	\caption{Factorial program with specification}
	\label{fig:factorial}
\end{figure}
Our function does not satisfy the specification. An overflow in the
32-bit Int leads to a negative result for the input 17, as Stainless
reports in Figure~\ref{fig:failed}. Changing the type from Int to
BigInt will result in a successful verification.

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{output1.png}
	\caption{Stainless output for the factorial program}
	\label{fig:failed}
\end{figure}

@@

\paragraph{Extracting the Relevant Code} The relevant code for the
addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
\texttt{NumberType.scala}.

\begin{figure}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}
  \caption{Extracted Code from NumberType.scala}
  \label{fig:numbertype}
\end{figure}

\begin{figure}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}
  \caption{Extracted Code from CurrencyUnits.scala}
  \label{fig:currencyunits}
\end{figure}

From those files we removed all code that is not needed for the
verification of our property.  For example, we removed all number
types except for Int64 (so Int32, UInt64, etc.) because they are not
used. We also removed the superclasses Factory and NetworkElement of
CurrencyUnit and Number, respectively, because the inherited members
are not used. Also, we removed all binary operations on Number that
are not used, like subtraction and multiplication. The extracted code
is shown in Figure~\ref{fig:numbertype} and
Figure~\ref{fig:currencyunits}.

@@

\paragraph{Transforming the Code.}

When we run Stainless on this code (without any properties to prove),
it throws the following errors: describe errors: no support for
  abstract types, unsupported arguments for the BigInt constructor,
  unsupported inheritance for objects.

\todo{...}

why fix the preconditions:
\begin{itemize}
\item all green: new errors stand out more
\item at runtime, errors are caught earlier, which is preferrable
\end{itemize}

@@

Finally, everything is green and correctly verified. 
\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{result_output}
	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
\end{figure}

@@

% \section{Extracted Code}
% \label{sec:extracted}

% \subsection{NumberType.scala}
% \lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}

% \subsection{CurrencyUnits.scala}
% \lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}

@@

\begin{lstlisting}[style=stainless]
[ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
            def toLong: Long = toBigInt.bigInteger.longValueExact()
                              ^^^^^^^^^^^^^^^^^^^
\end{lstlisting}
\todo{stainless error emssage missing}

@@

\todo{kai here we do not know if it changes semantic. ramon: oh, we don't?}

@@

Stainless does not support inner classes in a case object. Bitcoin-s
uses this a lot to separate the class \todo{ramon: i don't understand} from its implementation.

@@

% \section{Verified Code}
% \label{sec:verified}

% \subsection{NumberType.scala}
% \lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/number/NumberType.scala}

% \subsection{CurrencyUnits.scala}
% \lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}

>>>>

\begin{figure}
\begin{lstlisting}[style=scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}
	\caption{Factorial function with specification}
	\label{fig:factorial}
\end{figure}
Our function does not satisfy the specification. An overflow in the
32-bit Int leads to a negative result for the input 17, as Stainless
reports in Figure~\ref{fig:failed}. Changing the type from Int to
BigInt will result in a successful verification.

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{output1.png}
	\caption{Stainless output for the factorial function}
	\label{fig:failed}
\end{figure}

@@

\begin{figure}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/extracted/number/NumberType.scala}
  \caption{Extracted Code from NumberType.scala}
  \label{fig:numbertype}
\end{figure}

\begin{figure}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/extracted/currency/CurrencyUnits.scala}
  \caption{Extracted Code from CurrencyUnits.scala}
  \label{fig:currencyunits}
\end{figure}

\paragraph{Extracting the Relevant Code} The relevant code for the
addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
\texttt{NumberType.scala}. From those files we removed the majority of
the code because it is not needed for the verification of our
property.  For example, we removed all number types except for Int64
(so Int32, UInt64, etc.) because they are not used. We also removed
the superclasses Factory and NetworkElement of CurrencyUnit and
Number, respectively, because the inherited members are not
used. Also, we removed all binary operations on Number that are not
used, like subtraction and multiplication. The extracted code is shown
in Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.

@@

\paragraph{Transforming the Code.} Here we see how to transform the
bitcoin-s code into the Scala fragment supported by Stainless.  All
transformations are equivalent in the sense that if the
addition-with-zero property holds for the transformed code, then it
also holds for the code before the transformation. Here is the list of
features used by the extracted code but unsupported by Stainless, and
how to transform the code:

\begin{enumerate}
\item Inheriting Objects. Here, we can just turn the objects into case
  objects by literally just changing the word \texttt{object} into
  \texttt{case object} on lines 65 and 33 in the two respective files.

\item Abstract Type Members.

\item{Non-Literal BigInt Constructor Argument}

\item{Self-Reference in Type Parameter Bound}

\item{Missing Member \texttt{bigInteger} in BigInt}

\item{Type Member}

\item{Missing Bitwise-And Method on BigInt}

\item{Inner Class in Case Object}

Stainless does not support inner classes in a case object. Bitcoin-s
uses this a lot to separate the class from its implementation.

\item{Message Parameter in Require}

\item{Missing Implicit Long to BigInt Conversion}

\item{Missing BigInt Constructor with Long Argument}

\begin{lstlisting}[style=scala]
  lazy val min = Int64(BigInt("-9223372036854775808"))
  lazy val max = Int64(BigInt("9223372036854775807"))
}
\end{lstlisting}
\end{enumerate}

@@

Finally, everything is green and correctly verified. 
\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{result_output}
	\caption{Stainless output for the transformed code}
\end{figure}

@@

@@

\begin{lstlisting}[style=stainless]
[ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
            def toLong: Long = toBigInt.bigInteger.longValueExact()
                              ^^^^^^^^^^^^^^^^^^^
\end{lstlisting}

@@

@@

Stainless does not support inner classes in a case object. Bitcoin-s
uses this a lot to separate the class from its implementation.

@@
